---
title: "Inference of population structure and genetic differentiation"
author: "Danilo Lyra (Rothamsted Research, CAS)"
date: "May, 2020"
output:
  html_document:
  smart: no
  pdf_document: default
---

# Background
We detected the genomic diversity structure of the population at the gene level using the variants from QC scenario A. First, we applied a principal component (PC) analysis using the SNPRelate R package (snpgdsPCA function) (Zheng et al., 2012). Second, we applied a discriminant analysis of principal components (DAPCs) using the adegenet R package (Jombart et al., 2010). Group clustering used was inferred by Molero et al. (2019). The contributions (loadings) of each gene variant were estimated using the loadingplot function. Finally, a neighbor-joining tree (NJT) was generated based on the modified Euclidean distance using the ape R package (Paradis et al., 2004) and the pairwise genetic distance between populations (Fst) was calculated following Weir and Cockerham (1984) in SNPRelate R package. The genome-wide marker data, generated using the 35K Affymetrix Axiom® HD wheat SNP array (Allen et al., 2017), was used only for the DAPCs.


```{r packages, echo=T, message=FALSE, results=FALSE, warning = FALSE}
setwd("C:/Users/rnaly/OneDrive/Desktop/Data_archive_exome")

library(data.table) # fread
library(SNPRelate) #PCA and Fst
library(ggplot2) #plot
library(grid)
library(plyr)
library(adegenet) #DAPC
library(ape) #NJT
library(reshape) #melt
library(snpReady) #popgen
library(LDheatmap) #LD block
library(snpStats) #inout LDblock

load("Exome_snps_maf_1%.RData")

```

## Principal Component Analysis (PCA)

```{r PCA, echo=T}

#Load data
SNP_line=data.matrix(M)
id.sample = rownames(M)
id.snp = colnames(M)

M1=as.matrix(M)

# Create a gds file
snpgdsCreateGeno("ttest.gds", genmat = M1,
                 sample.id = id.sample, snp.id = id.snp,
                 snpfirstdim=F)
#Open the GDS file
(genofile <- snpgdsOpen("ttest.gds"))

## PCA
pca <- snpgdsPCA(genofile, eigen.cnt = 149)

eigenvect = pca$eigenvect
eigenval = pca$eigenval
pc.percent = pca$varprop*100
head(round(pc.percent, 2)) # variance pca

## Variance of PCA
barplot(eigenval,  
        main = "Variances",
        xlab = "Principal Components",
        ylab = "Percentage of variances",
        col ="steelblue")

```



## PCA plot

```{r PCA plot, echo=T}

myCol=c("#99C794","#6699CC")
PC_values=read.table("PC_exome.txt", header=T)

ggplot(PC_values, aes(x=PC1, y=PC2, colour = Group, fill = Group)) +
  scale_size_area() +
  geom_point(size = 9, pch = 21, stroke = 1, alpha=0.9, color="black") +
  
  geom_vline(xintercept = 0, lty="dotted",color="Black",size=0.8) +    # line in X-Axis
  geom_hline(yintercept = 0, lty="dotted",color="Black",size=0.8) +
  
  ## lines in regression
  scale_color_manual(values = c("#99C794", # Elite
                                "#6699CC" # Exotic
  )) +
  
  scale_fill_manual(values = c("#99C794", # Elite
                               "#6699CC" # Exotic
  )) +

  # identify and insert names in graph; hjust change the position
  geom_text(aes(label = Genotype), hjust=0.6, vjust=1.3, size=4) +

theme(legend.position="none", #bottom
      legend.title=element_text(size=13,color="black"),
      legend.text=element_text(size=12,color="black"),
      axis.text=element_text(color="black",size=14),
      axis.title=element_text(size=14,color="black",face="bold"), #axis size 
      strip.text= element_text(size = 14, colour = "black", face ="bold"),
      axis.text.x = element_text(angle = 90, size=14,vjust=0.01),
      panel.border = element_rect(linetype = "solid", size = 1, fill = NA, colour = "black"), ##all panel
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.ticks = element_line(size = 1, colour = "black"), # espessura dos pausinhos dos eixos
      axis.ticks.length = unit(0.2, "cm")
) 


```


## Fst analysis - exome data

```{r Fst, echo=T}

# Get sample id
sample.id <- read.gdsn(index.gdsn(genofile, "sample.id"))

#Pop group
Pop=read.table("Pop_group.txt", h=T)
pop_code=Pop$Group

flag <- pop_code %in% c("Elite", "Exotic")
samp.sel <- sample.id[flag]
pop.sel <- pop_code[flag]
Fst=snpgdsFst(genofile, 
          sample.id=samp.sel, 
          population=as.factor(pop.sel),
          method="W&C84")
Fst$Fst #Fst value

```

## Neighbor-joining tree (NJT) - Euclidean distance 

```{r NJT, echo=T}

load("all.data_scenarioA.RData")

x <- new("genlight", M) # conversion
x = as.matrix(x, parallel=FALSE)

tre <- nj(dist(as.matrix(x)))  #method =  "euclidean"


group1=read.table("group.txt", header=T)
group1$Color=as.character(group1$Color)
plot(tre, typ="fan", cex=0.7, show.tip=T)
plot(tre, typ="fan", show.tip=F)
tiplabels(pch=20, col=group1$Color, cex=4)


```


## Discriminant analysis of principal components (DAPCs) - 35k SNP Chip

```{r DAPC_35K, echo=T}
load("Snps_35K.RData")
x <- new("genlight", M_35K) # conversion
x=as.matrix(x, parallel=FALSE)

group=read.table("grp_35K.txt", h=T)
group=factor(group$Group)

set.seed(5)

##### DAPC
dapc1 <- dapc(x, group, parallel=FALSE, 
              n.pca=100, 
              n.da=1)

###################
#### Scatter DAPC
###################
myCol=c("#99C794","#6699CC")
scatter(dapc1, grp=dapc1$grp, posi.da="bottomright", bg="white",
        pch=17:22, 
        cstar=1,  #line to connect dots
        col=myCol, 
        scree.pca=T,
        solid=.9, #intensity of dot`s colour`
        cell=1.5, 
        cex=1.5, #size of dots
        posi.pca="topright", leg=T, 
        posi.leg="topleft",
        #posi.da="topleft",
        scree.da=T)

```

## DAPC - Exome-capture data

```{r DAPC_exome, echo=T}

load("all.data_scenarioA.RData")

x <- new("genlight", M) # conversion
x = as.matrix(x, parallel=FALSE)

group=read.table("Pop_group.txt", header=T)
group=factor(group$Group)

##### DAPC
set.seed(5)
dapc1 <- dapc(x, group, parallel=FALSE, 
              n.pca=50, 
              n.da=1)

## SNP contributions
loadingplot(dapc1$var.contr,thres=0.01)

#################
#### Scatter DAPC
#################
myCol=c("#99C794","#6699CC")

scatter(dapc1, grp=dapc1$grp, posi.da="bottomright", bg="white",
        pch=17:22, 
        cstar=1,  #line to connect dots
        col=myCol, 
        scree.pca=T,
        solid=.9, #intensity of dot`s colour`
        cell=1.5, 
        cex=1.5, #size of dots
        posi.pca="topright", leg=T, 
        posi.leg="topleft",
        #posi.da="topleft",
        scree.da=T)

```


## Polymorphic information content (PIC) and minor allele frequency (MAF) using the popgen function

```{r PIC, echo=T}

load("all.data_scenarioA.RData")

gen_par=popgen(M)
##snps info
snps_info=gen_par$whole$Markers
snps_info$gene=map_genes1$Gene_ID
## genotype info
genotype_info=gen_par$whole$Genotypes
##pop info
pop_info=gen_par$whole$Population

##################
#### Mean by Genes
##################
melted <- melt(snps_info, id.vars=c("gene"))

stats_gene=ddply(melted, c("gene", "variable"), summarise,
                 mean = mean(value), sd = sd(value),
                 sem = sd(value)/sqrt(length(value)))
stats_gene[1:10,]

```

## Linkage disequilibrium (LD) block - Subpopulations

```{r LD_group, echo=T}

load("MAP_Group.RData")
load("M_ElitevsExotic.RData")

M_ELITE=as.matrix(M_ELITE)
M_EXOTIC=as.matrix(M_EXOTIC)

#Elite
gdat<-as(M_ELITE,"SnpMatrix")
MyHeatmap <- LDheatmap(gdat, 
                       genetic.distances = MAP_ELITE$Pos, #map are the same for both elite and exotic
                       color = heat.colors(20),   #grey.colors(20)  heat.colors(20)
                       flip=TRUE) #color="blueToRed"

# Exotic
gdat<-as(M_EXOTIC,"SnpMatrix")

MyHeatmap <- LDheatmap(gdat, 
                       genetic.distances = MAP_EXOTIC$Pos, #map are the same for both elite and exotic
                       color = heat.colors(20),   #grey.colors(20)
                       flip=TRUE) #color="blueToRed"



```

## Linkage disequilibrium (LD) block - Gene

```{r LD, echo=T}

# Gene TraesCS1A02G064800
load("M_TraesCS1A02G064800.RData")
load("MAP_TraesCS1A02G064800.RData")
M_GENE=as.matrix(M_GENE)

gdat<-as(M_GENE,"SnpMatrix")

MyHeatmap <- LDheatmap(gdat, 
                       genetic.distances = MAP_GENE$Pos,
                       color = grey.colors(20),
                       flip=TRUE) #color="blueToRed"

```





