---
title: "**Screening for signature of selection at the gene level**"
author: "Danilo Lyra (Statistical & Quantitative Geneticist) - Rothamsted Research - CAS"
date: "May, 2020"
output:
  html_document:
  smart: no
  pdf_document: default
---

# Background
We evaluated the evidence of selection at the gene level by estimating the normalized ratio of nonsynonymous (missense, nonsense, and splicing) substitutions per synonymous site (ğœ” = dN/dS) using an optimized Poisson-based model (dNdScv) in the dndscv R package (Martincorena et al., 2017). Briefly, this model accounts for variation in mutation rates, sequence context, and full trinucleotide mutability. To estimate the mutation rate of a gene it uses a joint likelihood function combining local (synonymous substitutions in a gene) and global (negative binomial regression across genes) information to estimate the mutation rate of a gene. We used the buildref function to input the wheat reference genome (IWGSC RefSeq v1.0 annotation) from Ensembl Plants per chromosome. Global ğœ” estimates across all genes were estimated per chromosome. A global q-value â‰¤ 0.1 (without considering InDels) was used to identify statistically significant genes. A confidence interval (Î± = 0.95) was calculated per gene. Selection was measured as positive (ğœ” > 1), negative (ğœ” <1), and neutral (ğœ” =1) (Nielsen, 2005).


```{r packages, echo=T, message=FALSE, results=FALSE, warning = FALSE}
setwd("C:/Users/rnaly/OneDrive/Desktop/Data_archive_exome/Signature of selection")

library(dndscv) #Installation and tutorial: https://github.com/im3sanger/dndscv
library(data.table)

```

## Inputs for the buildref function 

```{r Data input, echo=T}
setwd("C:/Users/rnaly/OneDrive/Desktop/Data_archive_exome/Signature of selection")

#HiBAP wheat variants
mut_wheat=read.table("wheat_variants.txt", h=T)
mut_wheat$sampleID=as.character(mut_wheat$sampleID)
mut_wheat$chr=as.character(mut_wheat$chr)
mut_wheat$ref=as.character(mut_wheat$ref)
mut_wheat$mut=as.character(mut_wheat$mut)
head(mut_wheat)

#Table of transcripts - from biomart - Ensembl Plants
load("biomart_wheat.RData")
#http://htmlpreview.github.io/?http://github.com/im3sanger/dndscv/blob/master/vignettes/buildre#f.html
print(head(Biomart_wheat), digits = 3)

```

## dNdScv model

```{r Selection, echo=T}

#### CDS file
path_cds_table = file.path("C:", "Users", "rnaly", "OneDrive", "Desktop","DnDs_R",  "Biomart1.txt")

#############################################################################################
### Download the fasta files for each CHR in ftp://ftp.ensemblgenomes.org/pub/plants/release-46/fasta/triticum_aestivum/dna/
#############################################################################################

#Example of the Chr1A
path_genome_fasta_1A = file.path("C:", "Users", "rnaly", "OneDrive", "Desktop","DnDs_R",  "Triticum_aestivum.IWGSC.dna.chromosome.1A.fa")

#build Ref
buildref(cdsfile=path_cds_table, 
         genomefile=path_genome_fasta_1A, 
         outfile = "wheat_example_output_refcds_1A.rda", 
         excludechrs="MT")

#dndscv model to estimate dN/dS ratios
dndsout_1A = dndscv(mut_wheat, 
                    refdb="wheat_example_output_refcds_1A.rda", 
                    outmats=T)     ## estimate CI
                    #sm = "2r_3w" ## traditional transition/transversion (ts/tv) model
                    #cv=NULL)  #covariates


print(dndsout_1A$sel_cv) #Table of dN/dS ratios
(global_dnds = dndsout_1A$globaldnds) #Global dN/dS estimates
(annotmuts=dndsout_1A$annotmuts) #annotated table of coding mutations
(ci = geneci(dndsout_1A)) #confidence intervals for the dN/dS ratios

#Results from all chr are available in the Table S1
```

