---
title: "**Single variant association analysis**"
author: "Danilo Lyra (Statistical & Quantitative Geneticist) - Rothamsted Research - CAS"
date: "May, 2020"
output:
  html_document:
  smart: no
  pdf_document: default
---

# Background
Single variant association analysis was performed using a Mixed Linear Model (MLM) in GAPIT v3.0 R package (Lipka et al., 2012; Wang & Zhang, 2018) incorporating genomic kinship (K) matrix and the first three PCs (Q) to control for the confounding effects of cryptic relatedness and population structure (Yu et al., 2006). The default false discovery rate (FDR) (Benjamini & Hochberg, 1995) and Bonferroni multiple testing (Hochberg, 1988) was used to correct the genome-wide significance thresholds (α=0.05).


```{r packages, echo=T, message=FALSE, results=FALSE, warning = FALSE}
setwd("C:/Users/rnaly/OneDrive/Desktop/Data_archive_exome/Single-variant-analysis")

#devtools::install_github("jiabowang/GAPIT3",force=TRUE)
library(GAPIT3)
library(qqman) #QQ plot and Manhattan plot

source("http://www.zzlab.net/GAPIT/GAPIT.library.R")
source("http://www.zzlab.net/GAPIT/gapit_functions.txt")

```

## Data input

```{r Data input, echo=T}
##Load
load("Exome_snps_maf_1%.RData")

# Phenotype
my_pheno  <- read.table("y_data.txt", head = TRUE)
str(my_pheno)
#head(my_pheno)

# MAP file
{
  Z1map$chr[Z1map$chr=="1A"]=1
  Z1map$chr[Z1map$chr=="1B"]=2
  Z1map$chr[Z1map$chr=="1D"]=3
  Z1map$chr[Z1map$chr=="2A"]=4
  Z1map$chr[Z1map$chr=="2B"]=5
  Z1map$chr[Z1map$chr=="2D"]=6
  Z1map$chr[Z1map$chr=="3A"]=7
  Z1map$chr[Z1map$chr=="3B"]=8
  Z1map$chr[Z1map$chr=="3D"]=9
  Z1map$chr[Z1map$chr=="4A"]=10
  Z1map$chr[Z1map$chr=="4B"]=11
  Z1map$chr[Z1map$chr=="4D"]=12
  Z1map$chr[Z1map$chr=="5A"]=13
  Z1map$chr[Z1map$chr=="5B"]=14
  Z1map$chr[Z1map$chr=="5D"]=15
  Z1map$chr[Z1map$chr=="6A"]=16
  Z1map$chr[Z1map$chr=="6B"]=17
  Z1map$chr[Z1map$chr=="6D"]=18
  Z1map$chr[Z1map$chr=="7A"]=19
  Z1map$chr[Z1map$chr=="7B"]=20
  Z1map$chr[Z1map$chr=="7D"]=21
} # convert chr
Z1map$chr=as.numeric(Z1map$chr)

name_snp=rownames(Z1map)
map_gwas=Z1map[,c(1,2)]
map_gwas$marker=name_snp
map_gwas$marker=as.factor(map_gwas$marker)
map_gwas <- map_gwas[c("marker", "chr", "pos")] # reorder by column name

## GENO file
MM=as.data.frame(M)
name_id=rownames(MM)
myGD=cbind(name_id, MM)

```

## Single point scan - Mixed Linear Model (MLM)

```{r SVA_QCA, echo=T, message=FALSE, results=FALSE, warning = FALSE}

# Mixed Linear Model (MLM) 
myGAPIT <- GAPIT(
  #Y=my_pheno,        #all phenotypes
  Y=my_pheno[,c(1,10)], #selecting PH
  GD=myGD,
  GM=map_gwas,
  model="MLM",
  file.output=F,  #set to TRUE to save results in folder
  PCA.total=3
)


```

## Manhattan and QQ plot

```{r Plot_QCA, echo=T}

#Plant height (PH)
gap_gwas=myGAPIT$GWAS
gap_gwas=gap_gwas[order(gap_gwas$P.value),]
head(gap_gwas)

#QQplot
qq(gap_gwas$P.value)

#Manhattan plot
manhattan(gap_gwas, chr="Chromosome", bp="Position ", snp="SNP", p="P.value", 
          suggestiveline = -log10(0.05/749), #bonferroni correction
          genomewideline = F,
          annotatePval = 0.01)

```

